Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
added biblio

% Possible packages - uncomment to use
\usepackage{amsmath} % Needed for math stuff
%\usepackage{lastpage}sss% Finds last page
\usepackage{amssymb} % Needed for some math symbols
\usepackage{graphicx}% Needed for graphics
\usepackage[usenames,dvipsnames]{xcolor} % Needed for graphics and color
%\usepackage{setspace}sss% Needed to set single/double spacing
\usepackage{float} % Better placement of figures & tables
\usepackage{hyperref} % Can have actual links
\usepackage{mathpazo}
\usepackage[linesnumbered,vlined,ruled]{algorithm2e} % algorithm

@@

Performance of parallel algorithms are usually measured in terms of 3 different metrics: speedup, strong scalling efficiency and weak scaling efficiency.  Strong and weak scaling are common measures for the different ways a program can benefit from increased parallellism and more processors as well as the limitations that dictate the speedups. In each of the below subsections, $T(n)$ represents the computation of using $n$ processors and the same notation for speedup.

@@

The parareal algorithm, developed by Lions, Maday and Turinici in 2001
\cite{lions2001parareal}, is a generalized algorithm which allows for
parallelization in time. It does so by using a cheaper ($g_{\Delta t}$)---a
lower order or lesser resolution approximation first, which it then corrects by
evaluating segments with a finer method $g_{\textnormal{fine}}$. The correction
step from the approximation can be done in parallel. Multiple iterations can be
performed to achieve an error provably asymptotically equal to a full serial
computation of $g_{\textnormal{fine}}$. The exact number of iterations required
to achieve certain levels of convergence depends on the problem and must be
tuned top optimize for speedup for a given system. In summary, the algorithm is
$k$ repetitions of a correction process which requires running updated by a
finer method run in parallel for subsections.

@@

\textbf{Theorem}: The order of accuracy of the Parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
this using induction [Bal].

@@

Let us consider the ODE $du/dt = \lambda u$. Let $g_{\textnormal{fine}}(t_n,
u_n) = \bar{g_{\textnormal{fine}}}u_n$ and $g_{\Delta t}(t_n, u_n) =
\bar{g_{\Delta t}} u_n$. As shown by Staff et al. [cite], the parareal method
becomes

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$

@@

Varying Iterations:
]
Varying Quality Factor:
We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$

@@

\bibliographystyle{unsrt}
\bibliography{paper}

@@

Guilliame Bal paper

Stability of the Parareal Algorithm by Staff et al

>>>>

% Possible packages - uncomment to use
\usepackage{amsmath} % Needed for math stuff
%\usepackage{lastpage}sss% Finds last page
\usepackage{amssymb} % Needed for some math symbols
\usepackage{graphicx}% Needed for graphics
\usepackage[usenames,dvipsnames]{xcolor} % Needed for graphics and color
%\usepackage{setspace}sss% Needed to set single/double spacing
\usepackage{float} % Better placement of figures & tables
\usepackage{hyperref} % Can have actual links
\usepackage{mathpazo}
\usepackage[linesnumbered,vlined,ruled]{algorithm2e} % algorithm
\usepackage{microtype}

@@

Performance of parallel algorithms are usually measured in terms of 3 different
metrics: speedup, strong scalling efficiency and weak scaling efficiency.
Strong and weak scaling are common measures for the different ways a program can
benefit from increased parallellism and more processors as well as the
limitations that dictate the speedups~\cite{parallelscaling}. In each of the below subsections, $T(n)$ represents the computation of using $n$ processors and the same notation for speedup.

@@

The parareal algorithm, developed by Lions, Maday and Turinici in
2001~\cite{lions2001parareal}, is a generalized algorithm which allows for
parallelization in time. It does so by using a cheaper ($g_{\Delta t}$)---a
lower order or lesser resolution approximation first, which it then corrects by
evaluating segments with a finer method $g_{\textnormal{fine}}$. The correction
step from the approximation can be done in parallel. Multiple iterations can be
performed to achieve an error provably asymptotically equal to a full serial
computation of $g_{\textnormal{fine}}$. The exact number of iterations required
to achieve certain levels of convergence depends on the problem and must be
tuned top optimize for speedup for a given system. In summary, the algorithm is
$k$ repetitions of a correction process which requires running updated by a
finer method run in parallel for subsections.

@@

\textbf{Theorem}: The order of accuracy of the Parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
this using induction \cite{bal2005convergence}.

@@

Let us consider the ODE $du/dt = \lambda u$. Let $g_{\textnormal{fine}}(t_n,
u_n) = \bar{g_{\textnormal{fine}}}u_n$ and $g_{\Delta t}(t_n, u_n) =
\bar{g_{\Delta t}} u_n$. As shown by Staff et al. \cite{staff2005stability}, the parareal method becomes

@@

To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$

@@

Varying Iterations:

Varying Quality Factor:
We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$

@@

\nocite{*}
\bibliographystyle{unsrt}
\bibliography{paper}

@@
